\speculake: Hiding Malware in Speculative Execution

### 2018-08-03 paper/implementation.tex
impl pass

Instead, we
repeat the execution several times and look for the modal value over all
iterations. We measured the
error rate of our implementation as a function of how many redundant iterations
of the same step, and found that 10 redundant iterations resulted in 1 error every
million Turing steps, with the error rate dropping exponentially as iterations
increase. We choose 11 iterations as a conservative bound (error rate measured
to be 0), and computed 1 million Turing steps at a rate of 1351.75
steps per second.

@@

% TODO: needs a pass
%Constructing an emulator making use of the speculative primitive requires  
%a trade off in expresive capability versus speed. 
We have implemented our custom instruction set architecture---SPASM---as a model
using two pseudo-registers, and
6-bit instruction length which allows for a relatively direct programming model
in which structured values can be entered into memory locations before making
a systemcall.

@@

In SPASM we have implemented multiple example programs that we encrypted and
loaded into a \speculake payload, which decrypts and emulates SPASM instructions
only when the corresponding trigger program is running.
We first implemented a \textit{HelloWorld}
program that prints to \texttt{stdout}. Next, a \texttt{FizzBuzz} program that
demonstrates control flow and arithmetic operations while printing to stdout.
Finally, we implemented a \textit{ReverseShell} program that opens and connects to a socket
before duplicating I/O file descriptors and executing a local shell.
Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.

@@

To demonstrate a benign trigger application, we implemented an ExSpectre payload that
would trigger when running concurrently with OpenSSL.
% Could cut?
For these experiments, we disable
ASLR for simplification, but note that
branch predictors can also be used to determine ASLR offsets of co-resident applications,
and our attack adjusted accordingly~\cite{evtyushkin2016jump}.

We used \texttt{gdb} to run an instance of an OpenSSL server (version 1.0.1f),
and printed out every instruction executed and its address after a breakpoint on
the \texttt{SSL\_new} function. We then made a TLS connection to the
server,
which produced over 13~million instructions, including over 359,000 direct jumps
and 28,000 indirect jumps. We then searched for the longest repeated set of more than
28 indirect jumps that end with a unique jump (i.e. source and destination do
not occur in the previous 28+ indirect jumps).

We discovered a candidate that corresponds to code in OpenSSL's
\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times in our
handshake. This code is used during the TLS key exchange as the server computes
the ECDHE shared secret. We made a list of 31 source-destination address pairs
for these indirect jumps, and constructed a jump/ret chain to mimic the same
jump pattern in
our payload program. Our payload program mimicks the first 30 indirect jump
source/destination pairs, with a final jump going to a cache timing function
in our payload program. However, due to the prior pattern, this last jump is
frequently mis-speculated (about 3.5\% of the time), and instead goes to the
destination corresponding to the 31st jump in OpenSSL, which serves as our
speculative entry point.

@@

We verified that our payload program did not execute at the speculative entry
point when we ran
other programs that simply consumed CPU on the same core. In addition, when we used
Apache benchmark to create thousands of connections with a different cipher
suite (DHE-RSA-AES128-GCM-SHA256), we similarly received no speculation. This
could allow an adversary to use an obscure or uncommon (but still
supported) cipher suite to trigger a malicious \speculake payload program on a remote server.

>>>>

Instead, we
repeat the execution several times and look for the modal value over all
iterations. We measured the
error rate of our implementation as a function of how many redundant iterations
of the same step, and found that 10 redundant iterations resulted in 1 error every
million Turing steps, with the error rate dropping exponentially as iterations
increase. We choose 11 iterations as a conservative bound (error rate measured
to be 0), and computed 1 million Turing steps at a rate of 1351
steps per second.

@@

%Constructing an emulator making use of the speculative primitive requires  
%a trade off in expresive capability versus speed. 
We have implemented our custom instruction set architecture---SPASM---as a model
using two pseudo-registers, and
6-bit instruction length which allows for a relatively direct programming model
in which structured values can be entered into memory locations before making
a systemcall.

@@

In SPASM we have implemented multiple example programs that we encrypted and
loaded into a \speculake payload, which decrypts and emulates SPASM instructions
only when the corresponding trigger program is running.
We have implemented a \textit{HelloWorld}
program that prints to \texttt{stdout}, and a \texttt{FizzBuzz} program that
demonstrates control flow and arithmetic operations while printing to stdout.
Finally, we implemented a \textit{ReverseShell} program that opens and connects a TCP socket
to an attacker-chosen location
before executing a local shell and allowing the remote adversary
to issue shell commands on the victim machine.
Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.

@@

To demonstrate a benign trigger application, we implemented an ExSpectre payload that
would trigger when running concurrently with OpenSSL.
% Could cut?
We disable
ASLR for simplification, but note that
branch predictors can also be used to determine ASLR offsets of co-resident applications,
and our attack adjusted accordingly~\cite{evtyushkin2016jump}.

We used \texttt{gdb} to run an instance of an OpenSSL server (version 1.0.1f),
and printed out every instruction executed and its address after a breakpoint on
the \texttt{SSL\_new} function. We then made a TLS connection to the
server,
which produced over 13~million instructions, including over 359,000 direct jumps
and 28,000 indirect jumps. We then searched for the longest repeated set of more than
28 indirect jumps that ends with a unique jump (i.e. source and destination do
not occur in the previous 28+ indirect jumps).

We discovered a candidate that corresponds to code in OpenSSL's
\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times each
handshake. This code is used during the TLS key exchange as the server computes
the ECDHE shared secret. We made a list of 31 source-destination address pairs
for these indirect jumps, and constructed a jump/ret chain to mimic the same
jump pattern in
our payload program. Our payload program mimicks the first 30 indirect jump
source/destination pairs, with a final jump going to a cache timing function
in our payload program. However, due to the prior pattern, this last jump is
frequently mis-speculated (about 3.5\% of the time), and instead goes to the
destination corresponding to the 31st jump in OpenSSL, which serves as our
speculative entry point.

@@

We verified that our payload program did not execute at the speculative entry
point when we ran
other programs that simply consumed CPU on the same core. In addition, when we used
Apache benchmark to create thousands of connections with a different cipher
suite (DHE-RSA-AES128-GCM-SHA256), we similarly saw no speculation at the entry point. This
could allow an adversary to use an obscure or uncommon cipher suite to trigger a
malicious \speculake payload program on a remote server.
